<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Introduction to pangaear}
%\VignetteEncoding{UTF-8}
-->

```{r echo=FALSE}
knitr::opts_chunk$set(
  fig.path = "img/",
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

Introduction to pangaear
========================

`pangaear` is a data retrieval interface for the World Data Center PANGAEA (https://www.pangaea.de/). PANGAEA archieves published Earth & Environmental Science data under the following subjects: agriculture, atmosphere, biological classification, biosphere, chemistry, cryosphere, ecology, fisheries, geophysics, human dimensions, lakes & rives, land surface, lithosphere, oceans, and paleontology.

## Installation

If you've not installed it yet, install from CRAN:

```{r eval=FALSE}
install.packages("pangaear")
```

Or the development version:

```{r eval=FALSE}
devtools::install_github("ropensci/pangaear")
```

## Load pangaear

```{r}
library("pangaear")
```

## Search for data

`pg_search` is a thin wrapper around the GUI search interface on the page <https://www.pangaea.de/>. Everything you can do there, you can do here.

For example, query for the term 'water', with a bounding box, and return only three results.

```{r}
pg_search(query = 'water', bbox = c(-124.2, 41.8, -116.8, 46.1), count = 3)
```

The resulting `data.frame` has details about different studies, and you can use the DOIs (Digital Object Identifiers) to get data and metadata for any studies you're interested in.

### Another search option

There's another search option with the `pg_search_es` function. It is an interface to the Pangaea Elasticsearch interface. This provides a very flexible interface for search Pangaea data - though it is different from what you're used to with the Pangaea website. 

```{r}
(res <- pg_search_es())
```

The returned data.frame has a lot of columns. You can limit columns returned with the `source` parameter.

There are attributes on the data.frame that give you the total number of results found as well as the max score found. 

```{r}
attributes(res)
attr(res, "total")
attr(res, "max_score")
```

To get to the DOIs for each study, use 

```{r}
gsub("https://doi.org/", "", res$`_source.URI`)
```


## Get data

The function `pg_data` fetches datasets for studies by their DOIs.

```{r}
res <- pg_data(doi = '10.1594/PANGAEA.807580')
res[[1]]
```

Search for data then pass one or more DOIs to the `pg_data` function.

```{r}
res <- pg_search(query = 'water', bbox = c(-124.2, 41.8, -116.8, 46.1), count = 3)
pg_data(res$doi[3])[1:3]
```


## OAI-PMH metadata

[OAI-PMH](https://wiki.pangaea.de/wiki/OAI-PMH) is a standard protocol for serving metadata around objects, in this case datasets. If you are already familiar with OAI-PMH you are in luck as you can can use what you know here. If not familiar, it's relatively straight-forward. 

Note that you can't get data through these functions, rather only metadata about datasets.

### Identify the service

```{r}
pg_identify()
```

### List metadata formats

```{r}
pg_list_metadata_formats()
```

### List identifiers

```{r}
pg_list_identifiers(from = Sys.Date() - 2, until = Sys.Date())
```

### List sets

```{r}
pg_list_sets()
```

### List records

```{r}
pg_list_records(from = Sys.Date() - 1, until = Sys.Date())
```

### Get a record

```{r}
pg_get_record(identifier = "oai:pangaea.de:doi:10.1594/PANGAEA.788382")
```
